---
title: "Combine swemaps2 with data from Kolada"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  warning=FALSE,
  message=FALSE,
  out.width = "100%"
)
```

## Kolada

[Kolada](https://www.kolada.se/) is a data cooperation between Swedish municipalities and regions. Through their website you can find a rich variety of different statistics. Many of these come from Statistics Sweden (SCB) and some come from the regions themselves. To query this data you can use the excellent package [rKolada](https://lchansson.github.io/rKolada/).

Below we have a municipality map of Skåne that we want to fill with data from Kolada.

```{r}
library(tidyverse)
library(swemaps2)

municipality %>% 
  filter(ln_namn == "Skåne") %>% 
  ggplot() +
  geom_sf() +
  theme_swemap2()
```

[rKolada](https://lchansson.github.io/rKolada/) is a great package that is easy to use. Below we pull data from Kolada on how many housing projects that have been granted building approval.

```{r}
library(rKolada)

kpis <- get_kpi()

kpi_bygglov <- kpi_search(kpis, c("bostäder som beviljats bygglov"), column = "description")

kpi_bygglov_id <- kpi_extract_ids(kpi_bygglov)

municipalities <- get_municipality() %>% 
  municipality_extract_ids()

bygg_data <- get_values(
  kpi = kpi_bygglov_id,
  municipality =  municipalities
)
```

We can join the data by `municipality_id` and filter out one particular year: 

```{r}
skane_bygglov <- left_join(
  municipality, bygg_data, by = c("kn_kod" = "municipality_id")
  ) %>% 
  filter(year == 2019 & ln_namn == "Skåne") 

```

Next, we visualize the data using `geom_sf()`, we also use `geom_label_repel` from the package `ggrepel` to add some labels for municipalities.

```{r}
library(ggrepel)
ggplot(skane_bygglov, aes(fill = value)) +
  geom_sf() +
  theme_swemap2() +
  scale_fill_viridis_c(option = "magma")+
  labs(
    title = "Antal bostäder som beviljat bygglov",
    subtitle = "Antal bostäder som beviljats bygglov under två senast åren (T och T-1),\navser endast permanentboende, antal/1000 inv.",
    caption = "Källa: SKR och SCB genom Kolada",
    fill = "Antal\nbygglov"
  ) +
  geom_label_repel(
    data = head(skane_bygglov),
    aes(label = kn_namn, geometry = geometry),
    stat = "sf_coordinates",
    min.segment.length = 0,
    color = "darkgrey"
  )
```
